Stable quantum memories with few measurements

### 2017-04-21 IsingECC/manuscript_ising_ecc.tex
Small changes during meeting

 The choice of an Ohmic bath operationally sets the scaling of Lindblad operators that do not exchange energy with the environment \red{I don't get this last sentence. Is this all that this choice of bath does?}.  While detailed balance enforces the ratio of the rates of domain wall creation and annihilation to Boltzmann-like scaling (i.e., $\gamma(1)/\gamma(-1) = \exp(-1/T)$), the rate of error processes that do not involve energy exchange with the bath scales linearly with temperature.  In other words, $\gamma(0) \sim T$, which follows from taking the $\omega\rightarrow0$ limit of  Eq. \ref{eq:gammadef}.  Ultimately, this means that the hopping rate of domains walls is controlled by this choice of bath model \red{do you mean "...is controlled by temperature according to this choice of bath model"?}.  At finite temperatures, this introduces correlations into the patterns of errors that effect the system, and so it is no longer possible to talk about an ``independent error probability per site''.

@@

To calculate this probability, we proceed via Bayes rule: 
\begin{align}
\label{eq:bayes}
&P(d_1:d_2 | d_1^{x_1,t_1} \wedge d_2^{x_2,t_2}) = \\ \notag
&\frac{P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) P(d_1 : d_2)}{P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})}
\end{align}
The individual terms on the right hand side of equation \ref{eq:bayes} are straightforward to interpret.  $P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2)$ represents the probability that two measured defects would be at $(x_1,t_1)$ and $(x_2,t_2)$ given that they are indeed a pair.  This can be related to the probability that a one dimensional diffusion process with diffusion constant $D$ will have performed an excursion of length $|x_2-x_1|$ in time $|t_2-t_1|$, and takes the form,
\begin{align}
\label{eq:bayesformula}
P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) = 1 - {\rm erf}\bigg(\frac{|x_2-x_1|}{2\sqrt{D |t_2 - t_1|}}\bigg) 
\end{align}
This represents the probability that a random walk will not perform an excursion of distance $|x_2-x_1|$ in a time $t_2 - t_1$.  \red{This explanation confuses me. Why isn't this just a Gaussian?} For our analysis, we will choose $D\propto \gamma_0$.  The exact correspondence between $D$ and $\gamma_0$ depends on the details of the error correction algorithm itself, so, in practice, we treat the constant of proportionality as an empirically tuned parameter.

@@

 The two primary failure modes of the error correction procedure as temperature is made large are \red{Are there no failure modes for low temperature?} (1) defects escaping measurement patches, and (2) defects being paired incorrectly.  Failure mode 1 can be suppressed by working with large measurement patches, as the escape timescale is exponential in the number of sites within the measurement patch---i.e., it scales like $(1/\gamma_0)^{\rm{\#\ of\ sites\ on\ patch}}$.  Failure mode 2 can be suppressed to a degree, but at high enough temperature, will eventually happen with high probability, as is the case in the full-measurement version of the protocol.  If too many errors are introduced between measurement cycles, it becomes impossible to correct the system even in principle because the syndrome measurements can no longer be unambiguously decoded.  At sufficiently low temperature, however, with some assumptions, our algorithm can extend the lifetime of the system exponentially in system size.

@@

 In the following section, we provide numerical evidence that the lifetime of the Ising model in the presence of the protocol scales quasi-exponentially \red{why quasi?} with the number of measurement patches, as anticipated by the upper bound in Eq.\ref{eq:err_scaling}.

@@

 Broadly speaking, the algorithm is identical, but instead of having ``patches'' of measurement, there are measurement ``rails'', as indicated in \figref{fig:tc_fig}.  ``Centering'' of defects on rails amounts to shift-swaping defects into the center of the measurement rail \red{Cite your previous paper?}.  We conjecture that a sparse measurement strategy with randomly placed measurement patches of fixed diameter might exist for sufficiently large and sufficiently cold systems, \red{but we do not comment further on that here}.  The rail geometry of \figref{fig:tc_fig} is the simplest geometry that allows us to rely on the heuristic argument for a threshold sketched in Sec. \ref{sec:threshold} \red{rely on it to do what?}.

@@

More fundamentally, our algorithm can be understood as an entropy reduction scheme.  Configurations that give rise to errors in the encoded subspace are exponentially suppressed as system size is made larger.  This is in contrast with ``energetic'' suppression---that is, suppression \red{via cooling} or by widening the gap to excitations $\Delta$.  This tradeoff between entropic and energetic contributions is depicted in Figures \ref{fig:LifetimeVsTemperatureVsEnergy} and \ref{fig:LifetimeVsMeasurementFraction}.  Depending on the resource requirements of a particular architecture, the threshold temperature can be tuned either by engineering a larger gap, $\Delta$, or by changing the number of measurements used.  In practice, this will depend on the lowest effective temperature available, the maximum measurement rate, $\chi$, \red{Is $\chi$ used anywhere else?} as well as the practical difficulty of employing more measuring devices.

>>>>

 The main consequence of choosing an Ohmic bath is that it sets the amplitude of the excitation hopping process.  That is, $\gamma(0)$ is determined by the $\omega \rightarrow 0$ limit of the spectral density of the bath, and for the Ohmic bath taking the $\omega\rightarrow0$ limit of  Eq. \ref{eq:gammadef} yields $\gamma(0) \sim T$. 
Ultimately, this means that the hopping rate of domains walls is controlled by this choice of bath model. At finite temperatures, this introduces correlations into the patterns of errors that effect the system, and so it is no longer possible to talk about an ``independent error probability per site''. 
In contrast to the behavior of $\gamma(0)$, 
the other operationally important feature of the bath, the ratio of domain wall creation and annihilation rates, is set by detailed balance to Boltzmann-like scaling (i.e., $\gamma(1)/\gamma(-1) = \exp(-1/T)$), and is independent of the choice of bath spectrum.

@@

To calculate this probability, we proceed via Bayes rule: 
\begin{align}
\label{eq:bayes}
&P(d_1:d_2 | d_1^{x_1,t_1} \wedge d_2^{x_2,t_2}) = \\ \notag
&\frac{P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) P(d_1 : d_2)}{P(d_1^{x_1,t_1}\wedge d_2^{x_2,t_2})}
\end{align}
The individual terms on the right hand side of equation \ref{eq:bayes} are straightforward to interpret.  $P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2)$ represents the probability that two measured defects would be at $(x_1,t_1)$ and $(x_2,t_2)$ given that they are indeed a pair.  This can be related to the probability that a one dimensional diffusion process with diffusion constant $D$ will perform an excursion of distance $|x_2-x_1|$ or greater in a time $t_2 - t_1$; \textit{i.e.,} will perform an excursion that can measurement patches at $x_1$ and $x_2$. Explicitly,
\begin{align}
\label{eq:bayesformula}
P(d_1^{x_1,t_1} \wedge d_2^{x_2,t_2} | d_1:d_2) &= \red{\int_{x_1}^{x_2} ds (.)} \nonumber \\ 
&= 1 - {\rm erf}\bigg(\frac{|x_2-x_1|}{2\sqrt{D |t_2 - t_1|}}\bigg) 
\end{align}
For our analysis, we will choose $D\propto \gamma_0$.  The exact correspondence between $D$ and $\gamma_0$ depends on the details of the error correction algorithm itself, so, in practice, we treat the constant of proportionality as an empirically tuned parameter. \red{Mention approximation that the pair was created at the center.}

@@

 The two primary failure modes of the error correction procedure are (1) defects escaping measurement patches, and (2) defects being paired incorrectly. These failure modes become more harmful as temperature increases.  Failure mode 1 can be suppressed by working with large measurement patches, as the escape timescale is exponential in the number of sites within the measurement patch---i.e., it scales like $(1/\gamma_0)^{\rm{\#\ of\ sites\ on\ patch}}$.  Failure mode 2 can be suppressed to a degree, but at high enough temperature, will eventually happen with high probability, as is the case in the full-measurement version of the protocol.  If too many errors are introduced between measurement cycles, it becomes impossible to correct the system even in principle because the syndrome measurements can no longer be unambiguously decoded.  At sufficiently low temperature, however, with some assumptions, our algorithm can extend the lifetime of the system exponentially in system size.

@@

 In the following section, we provide numerical evidence that the lifetime of the Ising model in the presence of the protocol scales quasi-exponentially with the number of measurement patches, as anticipated by the upper bound in Eq.\ref{eq:err_scaling}.

@@

 Broadly speaking, the algorithm is identical, but instead of having ``patches'' of measurement, there are measurement ``rails'', as indicated in \figref{fig:tc_fig}.  ``Centering'' of defects on rails amounts to shift-swaping defects into the center of the measurement rail \red{Cite your previous paper?}.  We conjecture that a sparse measurement strategy with randomly placed measurement patches of fixed diameter might exist for sufficiently large and sufficiently cold systems.  The rail geometry of \figref{fig:tc_fig} is the simplest geometry that allows us to argue for a threshold temperature for the toric code, based on a generalization of the argument used for the Ising model in Sec. \ref{sec:threshold}. \red{Reverse these two sentences and soften language in the second sentence.}

@@

More fundamentally, our algorithm can be understood as an entropy reduction scheme.  Configurations that give rise to errors in the encoded subspace are exponentially suppressed as system size is made larger.  This is in contrast with ``energetic'' suppression---that is, suppression by widening the gap to excitations, $\Delta$ (or equivalently lowering the operating temperature).  This tradeoff between entropic and energetic contributions is depicted in Figures \ref{fig:LifetimeVsTemperatureVsEnergy} and \ref{fig:LifetimeVsMeasurementFraction}.  Depending on the resource requirements of a particular architecture, the threshold temperature can be tuned either by engineering a larger gap, $\Delta$, or by changing the number of measurements used.  In practice, this will depend on the lowest effective temperature available, the maximum measurement rate, $\chi$, \red{Is $\chi$ used anywhere else?} as well as the practical difficulty of employing more measuring devices.
